---
title: "Basic Shell Information"
author: "Niklas Hausmann"
date: "01/07/2020"
output:
  html_document: default
  pdf_document: default
---
# Introduction
This file describes the calculation of the basic data provided in Table 1 of the paper (i.e. the **total shell weight** and the **number of aperture measurements**).

First, we need to load the data that we want to work with:
```{r message=FALSE, warning=FALSE}
library(here)
library(tidyr)
library(dplyr)
Species <- read.csv(here("data/Species.csv"))
Aps <- read.csv(here("data/Apertures.csv"))
```


## Total shell weight


```{r echo=FALSE}
Species[1:5,1:5]
```

Some of the weights were measured using different scales. This was problematic, because some weights were too small to be registered by the scale we used in the field, and we had averaged their weight to 0.5 g. Following further analysis in the York laboratory with an improved scale, these weights turned out to be much lower, and we thus applied an average value of 0.25 g for those 0.5 g values.

```{r}
Species$Weight[Species$Weight==0.5] <- 0.25 #Change 0.5 weights to 0.25 weights 
```


Next, we restructured the dataframe to provide information for each sample as a whole. 
```{r}
Species <- Species %>%
  group_by(Site, Sample, Speciesname) %>% #this line and the next combines species that occur twice within one sample but which have not been added up yet. We could have done this step prior to importing the data but we had to adjust the weights first.
  summarise(Weight = sum(Weight)) %>%
  spread(Speciesname, Weight, fill = 0) #this lines is where we restructure the data

Species[1:6,c(1:2,4,6)]

```

We then want to combine those species-weights by sample into species-weights per site:
```{r}
Species <-
  aggregate(Species[, 3:97], by = list(Site = Species$Site), FUN = sum)

Species[1:6, 1:2]

```
However, we accidentally had the site JW1727 still split into its two columns A and B. 

```{r echo=FALSE}
Species[27:28,c(1,7,13)]
```

These need combining. 



```{r}
Species <- Species %>%
  separate(Site, sep = 6, into = "Site") %>% #here we simply remove the "_A" and "_B" part of the site name
  group_by(Site) %>% # then we can group the columns by that new value
  summarise_each(funs(sum)) # and group consisting of the two rows of JW1727 are summed up.

Species[27, c(1,7,13)]
```

Lastly, using the **Janitor** package and `adord_totals()`, we added the total value of all columns with weight values to the end of the dataframe. This is the data you find in Table 1 of the paper.
```{r echo=TRUE, message=FALSE, warning=FALSE}

library(janitor)

Species <- adorn_totals(Species, where = "col", name ="Total_weight")

Species %>% 
  select(Site,Total_weight) 

```

## Number of Apertures

The quite long 'Aps' dataframe can be summaries using the `summarySE.R` function, which we have to source first. You can read up more about this function (here)[http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/].


```{r}
Aps <- read.csv(here("data/Apertures.csv"))


source(here("/code/summarySE.R"))

Aps <-summarySE(na.omit(Aps), measurevar = "Aperture", groupvars = c("Site"))
head(Aps)

```



