---
title: "Aperture Statistics"
author: "Niklas Hausmann"
date: "13/03/2020"
output: html_document
---
# Introduction
This file explains how we ran the statisticla tests on the aperture size measurements.
It is based on the `Aps` dataframe, which we created in the **Basic Shell Information.R** file.
It is worth to redo these steps because we won't summarise it as much this time.
First, we need to source the *'Farasan Database.R'* file through:
```{r message=FALSE, warning=FALSE}
source('Farasan Database.R')
```

Then we extract the aperture measurements from the `Apertures` list:

```{r}
Aps <- data.frame()
f <- data.frame()

for (i in 1:length(Apertures))  {
  for (j in 1:length(Apertures[[i]]))  {
    if ("Strombus.fasciatus..............Born.1778..Aperture" %in% names(Apertures[[i]][[j]]) == FALSE) {
      next
    }
    f <-
      data.frame(c(as.character(names(Apertures[i]))), c(as.character(names(Apertures[[i]][j]))), Apertures[[i]][[j]][["Strombus.fasciatus..............Born.1778..Aperture"]])
    colnames(f) <- c('Site', 'Sample', 'Aperture')
    Aps <- rbind(Aps, f)
  }
}
rm(f, i, j, Areas, Sites, Species)


```
To get a simple list of each measurement grouped by site and bulk sample number:
```{r}
Aps[1:3,1:3]
```

# General aperture statistics by areas
To look at the general statistics for each areas, we added a column to `Aps` containing the respective bay areas and summarise the dataframe using the `summarySE` function:

```{r}
Aps$Area <- c(rep("Khur Maadi",4122), rep("Janaba West",6171),rep("Janaba East",5169))

source("summarySE.R")

Aps_area <-summarySE(na.omit(Aps), measurevar = "Aperture", groupvars = c("Area"))

```
```{r echo=FALSE}

Aps_area[1:3,1:5]

```

## One-way anova
Doing an one-way analysis of variance (anova) will show whether there are statistically significant differences between the means of each bay area.

```{r}
Aps_area_aov <- aov(Aperture ~ Area, data = Aps)
summary(Aps_area_aov)
```
## Post-hoc tukey test
We then applied a post-hoc tukey test, to understand between which areas the statistically significant differences lie.

```{r}
TukeyHSD(Aps_area_aov)
```

## Graph of size distributions by area (Fig. 6a)
To illustrate these differences, our paper used [raincloud plots](https://github.com/RainCloudPlots/RainCloudPlots), as well as other design-related packages, such as `viridis` and `cowplot`.

```{r message=FALSE, warning=FALSE,fig.width=5,fig.height=3}
library(ggplot2)
library(cowplot)
library(viridis)
source('R_rainclouds.R')

Fig6a <- ggplot(Aps, aes(
  x = Area,
  y = Aperture,
  fill = Area,
  colour = Area
)) +
  geom_flat_violin(
    position = position_nudge(x = +0.155, y = 0),
    adjust = 0.75,
    trim = FALSE,
    alpha = .99
  ) +
  geom_point(position = position_jitter(width = .1),
             size = 0.05,alpha=0.2
             
  ) +
  geom_boxplot(
    position = position_nudge(x = +0.155, y = 0),
    aes(x = Area, y = Aperture,colour=Area),
    outlier.shape = NA,
    alpha = 0.3,
    width = .1,
    colour = "BLACK",show.legend = FALSE
  ) +
  ylab('Aperture size') + xlab('') + coord_flip() + guides(fill = FALSE, colour = FALSE) +
  scale_color_manual(values=c(viridis_pal()(60)[10],viridis_pal()(6)[4],viridis_pal()(60)[59]))+
  scale_fill_manual(values=c(viridis_pal()(60)[10],viridis_pal()(6)[4],viridis_pal()(60)[59]))+
  theme_cowplot()+
  theme(axis.text.y.left = element_text(vjust=-2))+theme(axis.ticks.y=element_blank())

```
```{r}
Fig6a

```


# General aperture statistics by sites
To do the same tests between sites, we have to take similar steps, only this time we are also using the `Site` column to group aperture measurements.


```{r}


```

##Kruskal Wallis
The same data on site level reveals that there are variations in sample size, normal distributions and variance between the sites, which are all assumptions for the ANOVA test used above. We will thus switch to the non-parametric [Wilcoxon rank-sum test](https://en.wikipedia.org/wiki/Wilcoxon_signed-rank_test) and use the `pairwise.wilcox.test()` function

```{r message=FALSE, warning=FALSE}

PW_Aps <- pairwise.wilcox.test(Aps$Aperture, Aps$Site,  p.adjust.method = "bonferroni")
PW_Aps
```


## Graph of size distributions by site (Fig. 6b)
Similar to the graph above, we illustrate the site differences using [raincloud plots](https://github.com/RainCloudPlots/RainCloudPlots), as well as other design-related packages, such as `viridis` and `cowplot`. This time, the sites are coloured by bay area, however they are in the order of largst shells (top) to smallest shells (bottom). The general trend, however, remains and sites usually stick to their bay area.

```{r message=FALSE, warning=FALSE,fig.width=8,fig.height=8}
library(ggplot2)
library(cowplot)
library(viridis)
source('R_rainclouds.R')

 Fig6b <- ggplot(Aps, aes(
    x =   reorder(Site, Aperture, FUN = mean,order=TRUE),
    y = Aperture,
    fill = Area,
    colour = Area, order=Aperture
  )) +
    geom_flat_violin(
      position = position_nudge(x = +0.2, y = 0),
      adjust = 0.75,
      trim = FALSE,
      alpha = .99
    ) +
    geom_point(position = position_jitter(width = 0.1),
               size = 0.5,alpha=0.3
    ) +
    geom_boxplot(
      position = position_nudge(x = +0.2, y = 0),
      aes(x = Site, y = Aperture,colour=Area),
      outlier.shape = NA,
      alpha = 0.3,
      width = .1,
      colour = "BLACK"
    ) +
    ylab('Aperture size') + xlab("") + coord_flip() +guides(fill = guide_legend(reverse = TRUE),color = guide_legend(reverse = TRUE))+
    scale_color_manual(values=c(viridis_pal()(3)[1],viridis_pal()(6)[4],viridis_pal()(60)[59]))+
    scale_fill_manual(values=c(viridis_pal()(3)[1],viridis_pal()(6)[4],viridis_pal()(60)[59]))+    theme_cowplot()+ ggtitle("")+
    theme(axis.text.y.left = element_text(vjust=-0))+
    theme(axis.ticks.y=element_blank())+ 
    theme(legend.position="right")
  
  Fig6b
```


Finally, we want to integrate both raincloud plots (Areas and Sites) into one figure. For this we use the `patchwork` package. 



```{r message=FALSE, warning=FALSE,fig.width=10,fig.height=8}
library(patchwork)
Fig6 <- Fig6a+Fig6b

Fig6 + plot_annotation(
  title = 'Aperture sizes',
  subtitle = 'Differences between bay area on the left and differences between sites on the right')

setwd("~/Farasan-Shell-Data-master/results")#Remember to change the file path
ggsave("Figure_6.png", height = 200, width = 300, units = "mm")

```



# Time related changes in aperture sizes (Fig. 7)
First we will reload the `Aps` dataframe
```{r RerunAps}
Aps <- data.frame()
f <- data.frame()

for (i in 1:length(Apertures))  {
  for (j in 1:length(Apertures[[i]]))  {
    if ("Strombus.fasciatus..............Born.1778..Aperture" %in% names(Apertures[[i]][[j]]) == FALSE) {
      next
    }
    f <-
      data.frame(c(as.character(names(Apertures[i]))), c(as.character(names(Apertures[[i]][j]))), Apertures[[i]][[j]][["Strombus.fasciatus..............Born.1778..Aperture"]])
    colnames(f) <- c('Site', 'Sample', 'Aperture')
    Aps <- rbind(Aps, f)
  }
}
rm(f, i, j)
```

and add the Area information
```{r Sort Aps by Area}
Aps$Area <- c(rep("Khur Maadi",4122), rep("Janaba West",6171),rep("Janaba East",5169))

```

Then we will add the occupation times for each site. This step is done in the `SortApsByDate.R` script and essentially takes the minimum and maximum occupation period for each site and interpolates fictional dates for all bulk samples between the top and the bottom of the sample column. While this is not an ideal approach for analysing parallel changes between sites, they give us an idea of changes through time within individual sites and a rough grouping of data overall. 

```{r message=FALSE, warning=FALSE}
source('SortApsByDate.R')
```
```{r echo=FALSE}
ApsByDate[1,1:5]
```

To better illustrate trends within the dataset, we have averaged the aperture values for each bulk sample. That said, we did take into account the variability within each sample by scaling the size of each data point in the plot depending on each sample's variability (see further below in Fig. 7a).

```{r message=FALSE, warning=FALSE}
source("summarySE.R")
MeanAps <- summarySE(na.omit(ApsByDate[,1:5]), measurevar = c("Aperture"), groupvars=c("Sample"))
MeanYears <- summarySE(na.omit(ApsByDate[,1:5]), measurevar = c("Year"), groupvars=c("Sample"))
MeanApsByDate <- data.frame(MeanAps[,c(1,2,3,5)], Years_mean=MeanYears[,3])

MeanApsByDate <- merge(MeanApsByDate,ApsByDate,by="Sample")
MeanApsByDate <- MeanApsByDate[,c(8,6,1,2,3,4,5)]
rm(MeanAps,MeanYears)

MeanApsByDate <- distinct(MeanApsByDate)

head(MeanApsByDate)
```

## Figure 7a changes in time including all data
To show the changes in aperture size distribution through time, we plotted each bulk sample's mean aperture size as a point, that was itself larger or smaller depending on the variability of sizes within the specific bulk sample. We also overlayed this plot with the smoothed line using locally estimated scatterplot smoothing (LOESS).

```{r message=FALSE, warning=FALSE,fig.width=5,fig.height=5}
Fig7a <- ggplot(
  MeanApsByDate, 
  aes(x = Years_mean , y=Aperture_mean, size = sd)
) +scale_color_manual(values='steel')+
  geom_point(show.legend = TRUE, alpha = 0.4) + geom_smooth(method = "loess", span = 0.5, size = 0.5, alpha=0.1)+
  ylim(17.5,25)+ theme_bw()+ guides(size = FALSE)+ coord_flip()+theme(legend.position = c(0.8,0.3))+ylab('Mean aperture size')+xlab('Year cal BP')+
  scale_size(range = c(1.5, 6))+  ggtitle("Mollusc size change across Farasan")

Fig7a

```



## Figure 7b changes in time between bay areas
We are then looking at the aperture size distribution across time for each bay areas individually.

```{r message=FALSE, warning=FALSE,fig.width=5,fig.height=5}
library(viridis)
Fig7b <- ggplot(
  MeanApsByDate, 
  aes(x = Years_mean , y=Aperture_mean, size = sd, color=Area)
) +scale_color_manual(values=c(viridis_pal()(3)[1],viridis_pal()(6)[4],viridis_pal()(60)[59]))+
  geom_point(show.legend = TRUE, alpha = 0.4) + geom_smooth(method = "loess", span = 0.5, size = 0.5, alpha=0.1)+
  ylim(17.5,25)+ theme_bw()+ guides(size = FALSE)+ coord_flip()+theme(legend.position = c(0.8,0.3))+ylab('Mean aperture size')+xlab('Year cal BP')+
  scale_size(range = c(1.5, 6))+  ggtitle("Mollusc size change in different areas")


Fig7b

```




## Statistical test on trends between earlier and later periods

### Categorise aperture values by earlier and later periods within
To compare changes in time, which seem to be evident in Fig. 7b, we defined **Early** (6500--6000 calBP) and **Late** (5000--4800 cal BP) periods, into which the data from `ApsByDate` were grouped. Lastly these groups are combined into one dataframe (`Ttest.df`), which will be used to extract data for a two sample t-test in each bay area.

```{r}
EarlyAps <- ApsByDate[ApsByDate$Year<(-6000),]
EarlyAps <- EarlyAps[EarlyAps$Year>(-6500),]
EarlyAps$Period <- c("Early")

LateAps <- ApsByDate[ApsByDate$Year<(-4800),]
LateAps <- LateAps[LateAps$Year>(-5001),]
LateAps$Period <- c("Late")

Ttest.df <- rbind(EarlyAps,LateAps)

rm(EarlyAps,LateAps)

```
```{r echo=FALSE}
Ttest.df[c(1,1800),c(1,2,3,4,5,8)]
```

### Test for changes

To test changes in time for Janaba West:

```{r}
JW_Ttest.df <- subset(Ttest.df,Area=='Janaba West')
t.test(Aperture~Period, data=JW_Ttest.df, var.equal=TRUE)
```


To test changes in time for Janaba East:

```{r}
JE_Ttest.df <- subset(Ttest.df,Area=='Janaba East')
t.test(Aperture~Period, data=JE_Ttest.df, var.equal=TRUE)
```



## Temporal changes within individual sites
To see whether these changes persist on a site-level, we group the data by site.

```{r,fig.width=6,fig.height=6, message=FALSE, warning=FALSE}
Fig7c <- ggplot(
  MeanApsByDate, 
  aes(x = Years_mean, y=Aperture_mean, size = sd, color=Site)
) +scale_color_manual(values=c(viridis_pal()(60)[59],#The colours are a bit tricky so we had to specify them individually for best results
                               viridis_pal()(30)[28],
                               viridis_pal()(30)[25],
                               viridis_pal()(30)[22],
                               viridis_pal()(30)[19],
                               viridis_pal()(30)[17],
                               viridis_pal()(30)[10],
                               viridis_pal()(30)[6],
                               viridis_pal()(30)[4],
                               viridis_pal()(30)[1]))+
  geom_point(show.legend = TRUE, alpha = 0.4) +
  geom_smooth(method = "loess", span = 0.5, size = 0.5, alpha=0.1)+
  ylim(17.5,25)+ theme_bw()+ guides(size = FALSE)+ coord_flip()+theme(legend.position = c(0.8,0.3))+ylab('Mean aperture size')+xlab('Year cal BP')+
  scale_size(range = c(1.5, 6))+  ggtitle("Mollusc size change at different sites")

Fig7c


```

And we see that rather than a long-term decrease in size, we simply have changes in occupation of sites where shells are generally smaller. 

To combine all parts of Figure 6, we make use of `patchwork` again.


```{r, message=FALSE, warning=FALSE,fig.width=7*2.3,fig.height=7}


setwd("~/Farasan-Shell-Data-master/results")
Fig7 <- Fig7a+Fig7b+Fig7c
Fig7+plot_annotation(tag_levels = 'a', tag_suffix = ").")
ggsave("Figure_7.png", height = 180, width = 180*2.3, units = "mm")



```




